---
title: "SQL"
output: html_document
---

##Filtering Rows and Colomns

With SQL you can filter colomns with SELECT and rows with WHERE. Let's look at an example using the Lahman baseball database. First we need to load some packages. 

```{r warning=FALSE, message=FALSE}
library(Lahman)
library(sqldf)

```
#Example:

Suppose we want to see the homerun totals for the 1927 Yabkees. We could write the following.
```{r}
query<-"SELECT playerID, yearID, teamID, HR FROM Batting
WHERE teamID='NYA' and yearID=1927"
sqldf(query)

```

Moving on we can query out which NY Yankees players hit 40 or more homeruns with the following code:

```{r}
query<-"SELECT playerID,teamID,yearID,HR FROM Batting 
WHERE teamID='NYA' and HR>=40"
sqldf(query)

```
Adding the Strike Outs into the query we can further filter out which players hit over 40 homeruns AND had less than 60 strike outs with the following:
```{r}
query<-"SELECT playerID,teamID,yearID,SO,HR FROM Batting 
WHERE HR>=40 and SO<=60"
sqldf(query)

```
Finally we're filtering all the teams to just the Phillies, query the years in the 1970's (using a range), and finally extract the players with more than 30 homeruns.
```{r}
query<-"SELECT playerID,teamID,yearID,SO,HR FROM Batting 
WHERE yearID>1970 and yearID<1980 and  teamID='PHI' and HR>30"
sqldf(query)

```

##Ordering Data Output

With SQL you can also arrange the data by a certain order.By using the code order by, we are arranging the data based on HR in decending order

###Example

```{r}
query<-"SELECT playerID,teamID,yearID,HR FROM Batting
WHERE HR>50
ORDER BY HR DESC"
sqldf(query)

```
Here is another example of arranging the data by order:

```{r}
query<-"SELECT playerID,teamID,yearID,SO,AB FROM Batting
WHERE SO<10 and AB>=400
order by SO"
sqldf(query)

```

##Aggregation
Aggregation is essentially the "squishing" of all the data into 1 record.It allows for the ouput of the sum of all records and this can be done by using the code GROUP BY. To further narrow in on how the other columns should be grouped filtering can be done by using the HAVING command.

###Example:

Find Babe Ruth's career homerun total

```{r}
query<-"SELECT playerID, sum(HR) FROM Batting 
WHERE playerID='ruthba01'
GROUP BY playerID"
sqldf (query)

```
Find the career home run totals of all player, but limit the display to only those that hit 600 or more. Have the players with the highest total at the top.

```{r}
query<-"SELECT playerID,sum(HR) FROM Batting
GROUP BY playerID
HAVING sum(HR)>600
ORDER BY sum(HR)DESC"
sqldf(query)

```
Find the players with the highest homeruns average over their career. Limit the display to those who have an average of more than 30. Players with the highest average at the top.

```{r}
query<-"SELECT playerID,avg(HR) FROM Batting
GROUP BY playerID
Having avg(HR)>30
ORDER BY avg(HR)DESC"
sqldf(query)

```

##Joins
It allows tables to be joined together based on common data hence increasing the data output options. This is done through the INNER JOIN and ON commands. INNER JOIN indicates which tables are being joined together. ON indicates the common data that the tables will be joined by.

###Example:

List first name,last name, year, team ID and HR. Limit it to Babe Ruth

```{r}
query<-"SELECT nameFirst,nameLast,teamID,HR 
FROM Batting INNER JOIN Master
ON Batting.playerID=Master.playerID
WHERE Batting.playerID='ruthba01'"

sqldf(query)

```

Find all instances of players hitting more than 50 homeruns. Give first and last names, team ID, year ID and HRs. 

```{r}
query<-"SELECT nameFirst,nameLast,teamID,yearID,HR
FROM Batting INNER JOIN Master
ON Batting.playerID=Master.playerID
WHERE HR>50
ORDER BY HR DESC"

sqldf(query)

```

